GVN-Hoist: Hoisting Computations from Branches

### 2016-08-08 hoist.tex
scalar load store call

\subsubsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze
them for aliasing memory references. As long as all the operands are
available (which in SSA form we do, or may be there are chained
dependencies), the scalar can be hoisted. Care must be taken in case
of hoisting scalars too far, as that may increase register pressure
and result in spills. For example hoisting a scalar past a call. In
that case the call may result in save and restore of the register the
scalar may be defined to. In our current implementation we do not
hoist scalars past a call. Another way to mitigate this problem is be
to reinstantiate (rematerialize) the computation after a call (may be
as a different pass).

\subsubsection{Hoisting loads}
In order to value number stores, we value number (hash) the address from where
the value is to be loaded. Then we need to compute the hoisting point just like
we did for the scalars. At the hoisting point we compute if the operands to the
load (an address) is available. If it is not available we can try to compute the
addresss if that is possible.  Along the path from current position of the load
instruction through the hoiting point, we find if there are aliasing stores, we
find if any of the basic blocks are landing pads or targets of indirect
branches. In those cases we discard the candidate. Since the load increases the
register pressure by one (unless the operand to load becomes dead at the load),
we evaluate if it is beneficial to hoist the load.

\subsubsection{Hoisting stores}
In order to value number stores, we value number (hash) the address as well as
the value to be stored to that address. For the legality, we need to check the
dependency requirements just like hoisting loads. We need to check that the
operands of store are available at the hoisting point, that there are no
aliasing loads or store along the path from the current position through the
hoisting point and, none of the basic blocks along the path are landing pads.
Once all the legality checks are satisfied, it is a good candidate to be
hoisted.

\subsubsection{Hoisting calls}
Calls can be hoisted by value-numbering all the arguments. For safety
we have to check that there are no other side-effects between a call and
its hoisting point, just like loads and stores.

@@

\subsubsection{In the presence of calls}
Hoisting scalars across calls is tricky because it can increase the number of
spills. During the frame lowering of calls, the argument registers, in general,
the caller saved registers are saved because they might be modified by the
callee and after the call they are restored \cite{frame-lowering}. So before the
call, the register pressure is high because the number of available registers
are reduced by the number of caller saved registers. In that situation if a
computation is hoisted across the call, that would increase the total number of
registers required by one, thus contributing to the register pressure.

>>>>

\subsubsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available, the
scalar can be hoisted. Care must be taken in case of hoisting scalars too far,
as that may increase register pressure and result in spills. For example
hoisting a scalar past a call (Section-\ref{cost:across-calls}). In our current
implementation we hoist scalars past a call only when optimizing for code-side
(-Os). Another way to mitigate this problem is be to reinstantiate
(rematerialize) the computation after a call (may be as a different pass).

\subsubsection{Hoisting loads}
In order to value number stores, we value number (hash) the address from where
the value is to be loaded. Then we need to compute the hoisting point similar to
the hoisting of scalars. At the hoisting point the availability of operands to the
load (an address) is checked. If it is not available we try to compute the
addresss if that is possible.  Along the path, from current position of the load
instruction through the hoisting point, we find if there are aliasing stores, we
find if any of the basic blocks are landing pads or targets of indirect
branches. In those cases we discard the candidate. Since the load increases the
register pressure by one (unless the operand to load becomes dead at the load),
we also evaluate if it is beneficial to hoist the load.

\subsubsection{Hoisting stores}
In order to value number stores, we value number (hash) the address as well as
the value to be stored at that address. For the legality, we need to check the
dependency requirements similar to the hoisting of loads. We need to check that
the operands of the store instruction are available at the hoisting point, that
there are no aliasing loads or store along the path from the current position
through the hoisting point and, none of the basic blocks along the path have
their address taken. Once all the legality checks are satisfied, it is a good
candidate to be hoisted.

\subsubsection{Hoisting calls}
Calls can be hoisted by value-numbering all the arguments. For safety
we have to check that there are no other side-effects between a call and
its hoisting point, just like we did for loads and stores.

@@

\subsubsection{In the presence of calls}
\label{cost:across-calls}
Hoisting scalars across calls is tricky because it can increase the number of
spills. During the frame lowering of calls, the argument registers, in general,
the caller saved registers are saved because they might be modified by the
callee and after the call they are restored \cite{frame-lowering}. So before the
call, the register pressure is high because the number of available registers
are reduced by the number of caller saved registers. In that situation if a
computation is hoisted across the call, that would increase the total number of
registers required by one, thus contributing to the register pressure.
